### 2014-09-18 trygve_b_written_assignment1.tex
Started on Written assignment 1 in SPRÅK3501.

\documentclass[11pt, a4paper]{article}
\usepackage{mathptmx}
\usepackage{mathtools}
\usepackage{amssymb}
%\usepackage{amsfonts}
%\usepackage{amsmath}
\usepackage{fullpage}
\usepackage{enumerate}
\usepackage{amsthm}
\usepackage{remreset}
\usepackage[utf8]{inputenc}

\usepackage{lmodern}

\begin{document}

\title{Written assignment 1 in SPRÅK3501}
\author{Trygve Bærland}
\maketitle
\section*{First article:}
\cite{Cifani} is an article showing the main strategy I want to do.

\section*{Second article:}
\cite{Schochet}

\newpage
\bibliographystyle{plain}
\bibliography{project_bibliography.bib}
\end{document}