### 2014-10-15 stuff_somewhat_organised.tex
Fortsatt møkk.

\documentclass[11pt, a4paper]{article}
\usepackage{mathptmx}
\usepackage{mathtools}
\usepackage{amssymb}
%\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{fullpage}
\usepackage{enumerate}
\usepackage{amsthm}
\usepackage{remreset}
\usepackage[utf8x]{inputenc}
\usepackage[toc,page]{appendix}
\usepackage[numbers]{natbib}

\usepackage{lmodern}

\begin{document}
\title{Stuff that may well go into the project report}
\author{Trygve Bærland}
\maketitle
\abstract{For now, heavily reliant on \citep{vazquez2007porous}.}

% My text categories:
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{cor}{Corollary}[section]
\newtheorem{lemma}{Lemma}[section]

\theoremstyle{definition}
\newtheorem{mydef}{Definition}[section]
\newtheorem{example}{Example}[section]

%\theoremstyle{plain}
\newtheorem{obs}{Observation}
\newtheorem{rem}{Remark}

% Reset numbering for each section
\numberwithin{equation}{section}
\section{Basics}

Let $\Omega \subset \mathbb{R}^d$ be an open, bounded, and simply connected domain. Then the filtration, or general porous medium equation is

\begin{equation}
\label{GPME}
	\partial_t u = \Delta \Phi(u) + f,
\end{equation}

 where $u, f: \Omega \times (0,\infty) \to \mathbb{R}$, and $\phi: \mathbb{R} \to \mathbb{R}$. For now we assume that $\Phi$ is continuous and nondecreasing.

\subsection{Problem statement}
I will consider on the following Dirichlet problem:
\begin{equation}
\label{HDP}
	\begin{cases}
		\partial_t u = \Delta \Phi(u) + f, \quad \text{ in } Q_T \\
		u(x,0) = u_0(x), \quad \text{ in } \Omega \\
		u(x,t) = 0, \quad \text{ on } \partial\Omega \times (0,T).
	\end{cases}
\end{equation}
Here $Q_T = \Omega \times (0,T)$, where $T$ may very well be infinite.
\subsection{Weak solutions}
\begin{mydef}[Weak solutions]
A locally integrable function $u$ defined in $Q_T$ is said to be a \textbf{weak solution} of (\ref{HDP}) if
\begin{enumerate}[i)]
	\item $u \in L^1(Q_T)$ and $\Phi(u) \in L^1(0,T : W_0^{1,1}(\Omega))$\footnote{$f:\Omega \to \mathbb{R}$ is an element of $W^{k,p}(\Omega)$ if all weak derivatives with $|\alpha| \leq k$ exists and are in $L^p(\Omega)$. $W_0^{k,p}$ is the closure of $C^\infty_c(\Omega)$ in $W^{k,p}(\Omega)$.};
	\item $u$ satisfies
		\begin{equation}
		\label{weak}
		\iint_{Q_T} \{\nabla\Phi(u)\cdot \nabla\eta - u\partial_t \eta \}dxdt = \int_\Omega u_0(x)\eta(x,0)dx + \iint_{Q_T}f\eta dxdt
		\end{equation}
for any $\eta \in C^1(\overline{Q}_T)$ that vanishes on $\partial \Omega \times [0,T)$ and for $t=T$.
\end{enumerate}
\end{mydef}

\begin{obs}
With $u_0 \in L^1(\Omega)$ and $f\in L^1(Q_T)$, all the integrals in (\ref{weak}) make sense.
\end{obs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%		ENERGY SOLUTIONS:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Energy solutions}
We define  $\Psi$ as the primitive of $\Phi$ with $\Psi(0) = 0$, i.e.
\begin{equation*}
	\Psi(s) = \int_0^s \Phi(u) ds.
\end{equation*},
and we define the function space $L_\Psi(\Omega) = \{ u\in L^1(\Omega): \Psi(u)\in L^1(\Omega)\}$.
With this in mind we can define energy solutions in the following way:
\begin{mydef}[Energy solution]
We say that $u$ is a \textbf{weak energy solution} if it's a weak solution, and $u_0 \in L_\Psi(\Omega)$, and the energy inequality
\begin{equation}
\label{energy_ineq}
\iint_{Q_T} |\nabla \Phi(u)|^2 dxdt + \int_\Omega \Psi(u(x,T))dx \leq \int_\Omega \Psi(u_0)dx + \iint_{Q_T}f\Phi(u)dxdt
\end{equation}
holds.
\end{mydef}

\subsubsection{Derivation of energy estimate for classical solutions}
Suppose $u$ is a smooth solution of (\ref{HDP}). We proceed by multiplying the equation with $\Phi(u)$ to get
\begin{equation*}
\Phi(u)\partial_t u = \Phi(u)\Delta \Phi(u) + \Phi(u)f,
\end{equation*}
and we notice that $\Phi(u)\partial_t u = \partial_t \Psi(u)$. That means when we integrate over $Q_T$ we have
\begin{equation*}
\iint_{Q_T} \partial_t \Psi(u)dxdt = \int_\Omega \Psi(u(T))dx - \int_\Omega \Psi(u_0)dx.
\end{equation*}
On the diffusive term we can use integration by parts, so
\begin{equation*}
\iint_{Q_T}\Phi(u) \Delta \Phi(u) dxdt = -\iint_{Q_T} |\nabla \Phi(u)|^2dxdt + \int_0^T \int_{\partial\Omega} \Phi(u)\nabla \Phi(u) \cdot \hat{n}dS,
\end{equation*}
where $\hat{n}$ is the outward unit normal. By the Dirichlet condition
\begin{equation*}
\iint_{Q_T}\Phi(u) \Delta \Phi(u) dxdt = -\iint_{Q_T} |\nabla \Phi(u)|^2dxdt,
\end{equation*}
and notice that this step would have been valid for homogeneous Neumann condition as well. In toto 
\begin{equation}
\iint_{Q_T} |\nabla \Phi(u)|^2dxdt + \int_\Omega \Psi(u(T))dx = \int_\Omega \Psi(u_0)dx + \iint_{Q_T} f\Phi(u)dxdt.
\end{equation}

\setcounter{obs}{0}
\begin{obs}
In contrast to (\ref{energy_ineq}) we have equality for smooth solutions.
\end{obs}

\begin{obs}
We can go even further and deal with the last term on the right hand side: Assuming $f,\Phi(u) \in L^2(Q_T)$, then
\begin{equation*}
\iint_{Q_T} f\Phi(u)dxdt \leq \left(\iint_{Q_T} f^2 dxdt\right)^\frac{1}{2}\left(\iint_{Q_T} \Phi(u)^2 dxdt\right)^\frac{1}{2},
\end{equation*}
by Cauchy-Schwarz'. With the inequality $\sqrt{ab} \leq \frac{1}{4c}a + cb$ for $c>0$
\begin{equation*}
\iint_{Q_T} f\Phi(u)dxdt \leq \frac{1}{4c}\iint_{Q_T}f^2dxdt + c\iint_{Q_T}\Phi(u)^2dxdt.
\end{equation*}
Lastly, because of the homogeneous Dirichlet conditions, we use the Poincaré inequality and set $c$ so that
\begin{equation}
\label{using_Poincare}
\iint_{Q_T} f\Phi(u)dxdt \leq C\iint_{Q_T}f^2dxdt + \frac{1}{2}\iint_{Q_T}|\nabla \Phi(u)|^2dxdt,
\end{equation}
and the final inequality becomes
\begin{equation}
\frac{1}{2}\iint_{Q_T}|\nabla \Phi(u)|^2dxdt + \int_\Omega \Psi(u(T))dx \leq \int_\Omega \Psi(u_0)dx + C\iint_{Q_T}f^2dxdt.
\end{equation}
\end{obs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%					ESTIMATES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Some estimates}
Now we need to make some assumptions on $\Phi$ and $f$:
\begin{enumerate}[i)]
	\item The function $\Phi: \mathbb{R} \to \mathbb{R}$ is continuous, strictly increasing and we assume $\Phi(0) = 0$ without any loss of generality.
\end{enumerate}

For now the assumption that $\Phi'(u) > 0$ is kind of strong, but I hope to generalise (or find analogues for) most of the estimates to the case $\Phi'(u) \geq 0$. In that case (\ref{GPME}) changes from nodegenerate parabolic to degenerate parabolic. But hopes are high.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%					L^1-CONTRACTIVITY
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{$L^1$-contractivity}
Now we'll try to prove the following result:

\begin{proposition}
For two weak energy solutions $u$ and $\hat{u}$ with initial data $u_0$, $\hat{u}_0$ and forcing terms $f$, $\hat{f}$ respectively, we have for every $t > \tau \geq 0$
\begin{equation}
\label{l1_con}
||(u(t) - \hat{u}(t))_+||_1 \leq ||(u(\tau) - \hat{u}(\tau))_+||_1 + \int_\tau^t ||(f(s) - \hat{f}(s))_+||_1ds. 
\end{equation}
\end{proposition}

The proof will be done by establishing the result for smooth solutions of (\ref{HDP}). For such a solution to exist we must require that $\Phi'(u) > 0$, as well as some smoothness requirements on $f$ and $\partial\Omega \times (0,T)$ (I will check these smoothness assumptions). Then approximating the problem by smooth data, for which the $L^1$-contractivity holds, and pass to the limit.

\subsubsection{$L^1$-contractivity for smooth solutions}
Let $p: \mathbb{R} \to \mathbb{R}$ be a smooth function satisfying $p(s) = 0$ for $s \leq 0$, $p'(s) \geq 0$ for all $s\in \mathbb{R}$, and $0 \leq p(s) \leq 1$ for $s > 0$. Define now $w = \Phi(u) - \Phi(\hat{u})$, and multiply $p(w)$ with the difference of (\ref{GPME}) for $u$ and $\hat{u}$ to get
\begin{equation*}
\partial_t(u-\hat{u})p(w) = p(w)\Delta w + p(w)f.
\end{equation*}
Integrating over $\Omega$, and using integration by parts on the first part on the right hand side yields
\begin{align*}
\int_{\Omega}\partial_t(u-\hat{u})p(w)dx &= -\int_{\Omega}p'(w)|\nabla w|^2dx + \int_{\Omega}p(w)(f-\hat{f}) dx \\
	&\leq \int_{\Omega}p(w)(f-\hat{f}) dx \\
	&\leq \int_{\Omega}(f-\hat{f})_+ dx,
\end{align*}
where the last step follows from the fact that $0 \leq p(w) \leq 1$.
Now, let $p$ converge to the Heaviside function $H(w)$. Then, it is worth noticing that $H(w) = H(u-\hat{u})$, because of the strict monotonicity of $\Phi$. Also see that $\partial_t(u-\hat{u})_+ = \partial_t(u-\hat{u})H(u-\hat{u})$. Thus we get that
\begin{equation*}
\partial_t \int_{\Omega}(u-\hat{u})_+dx \leq \int_{\Omega}(f-\hat{f})_+ dx,
\end{equation*}
and integrating in time from $\tau$ to $t$ yields
\begin{equation*}
\int_{\Omega}(u(t)-\hat{u}(t))_+dx \leq \int_{\Omega}(u(\tau)-\hat{u}(\tau))_+dx + \int_\tau^t\int_{\Omega}(f-\hat{f})_+dxdt.
\end{equation*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%					UNIQUENESS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Uniqueness}
\begin{theorem}[Uniqueness]
	Assuming in addition that $u \in L^2(Q_T)$ and $\Phi(u) \in L^2(0,T : H_0^1(\Omega))$, (\ref{HDP}) has at most one weak solution.
\end{theorem}
\begin{proof}
This is a pretty neat proof using a smart choice of $\eta$. So let's assume $u_1$ and $u_2$ are weak solutions. Then, when putting $u_1$ and $u_2$ into (\ref{weak}) and subtracting one from the other, we get
\begin{equation}
\label{weak_diff}
\iint_{Q_T}\nabla(\Phi(u_2) - \Phi(u_1)) \cdot \nabla \eta - (u_2 - u_1)\partial_t \eta dxdt = 0.
\end{equation}
Now the inspired move\footnote{Of course, this isn't my proof so that sort of horn-tooting is in good taste.} is choosing
\begin{equation*}
\eta(x,t) = \begin{cases}
		\int_t^T(\Phi(u_2) - \Phi(u_1))ds, \quad &\text{ if } 0 < t < T \\
		0 \quad &\text{ if } t\geq T.
	\end{cases}
\end{equation*}
We then have
\begin{equation*}
\begin{cases}
	\partial_t \eta &= -(\Phi(u_2) - \Phi(u_1) ) \\
	\nabla \eta &= \int_t^T\nabla (\Phi(u_2) - \Phi(u_1))ds,
\end{cases}
\end{equation*}
where the time derivation is valid using the First Fundamentel Theorem of Calculus and that $\Phi(u_i) \in L^1(0,T : W_0^{1,1}(\Omega))$ by the definition of weak soultion. For the validity of the spatial derivative, confer appendix \ref{sec:derivative}.

When putting these derivatives into the above equality we end up with
\begin{equation}
\label{unique}
\begin{aligned}
	&\iint_{Q_T}(\Phi(u_2) - \Phi(u_1))(u_2 - u_1)dxdt \\ 
	&+ \iint_{Q_T}\nabla(\Phi(u_2(t)) - \Phi(u_1(t))) \cdot \left(\int_t^T 					\nabla(\Phi(u_2(s)) - \Phi(u_1(s))ds\right)dtdx = 0.
\end{aligned}
\end{equation}
Let's take these in order:
\begin{enumerate}[i)]
	\item In the first integral we notice that $u_2 > u_1 \Rightarrow \Phi(u_2) \geq \Phi(u_1)$ and vice versa, so this integral is nonnegative.

	\item For the second integral it is helpful to first only consider integration of the temporal dimensions, leaving us with
	\begin{align*}
	&\int_0^T\int_t^T \nabla(\Phi(u_2(t)) - \Phi(u_1(t)))\cdot \nabla(\Phi(u_2(s))-\Phi(u_1(s)))dsdt \\
	&= \int_0^T\int_0^s\nabla(\Phi(u_2(t)) - \Phi(u_1(t)))\cdot \nabla(\Phi(u_2(s))-\Phi(u_1(s)))dtds.
	\end{align*}
	Due to the symmetry about the line $s=t$, we have
	\begin{align*}
	&\int_0^T\int_0^s\nabla(\Phi(u_2(t)) - \Phi(u_1(t)))\cdot \nabla(\Phi(u_2(s))-\Phi(u_1(s)))dtds \\
	&= \frac{1}{2}\int_0^T \int_0^T \nabla(\Phi(u_2(t)) - \Phi(u_1(t)))\cdot \nabla(\Phi(u_2(s))-\Phi(u_1(s)))dtds \\
	&= \frac{1}{2}\left( \int_0^T \nabla(\Phi(u_2)-\Phi(u_1))dt\right)^2,
	\end{align*}
	which is also nonnegative, and therefore the second integral in (\ref{unique}) must also be nonnegative.
\end{enumerate}
From this we may conclude that both integral terms in (\ref{unique}) are zero, which in turn implies, by the first integral that $(\Phi(u_2)-\Phi(u_1))(u_2 - u_1) = 0$ a.e. in $Q_T$. Actually, we have that $\Phi(u_2) - \Phi(u_1) = 0$ a.e., because wherever $u_2 = u_1$ it is trivially true, and wherever $ u_2 \neq u_1$ it holds from the first integral being zero. Thus, $ \nabla (\Phi(u_2) - \Phi(u_1) ) = 0$ a.e. and putting this into (\ref{weak_diff}) we get that
\begin{equation*}
\iint_{Q_T}(u_2 - u_1)\partial_t \eta dx dt = 0.
\end{equation*}
This should hold for any test function $\eta$, and so the weak derivative of $u_2 - u_1$ with respect to time is zero. Because $u_2$ and $u_1$ have identical initial condition $u_0$ we conclude that $u_2 = u_1$.
\end{proof}

\setcounter{obs}{0}
\begin{obs}
Provided $\iint_{Q_T} f\Phi(u)dxdt$ in (\ref{energy_ineq}) makes sense, $\Phi(u) \in L^2(0,T: H_0^1(\Omega))$ for weak energy solutions. If we can put on a condition on $u$ or any of the data to ensure that $u \in L^2(Q_T)$, we would have a condition under which weak energy solutions are unique. Hopefully, no such condition is required, but will perhaps follow from some other estimate.
\end{obs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%			EXISTENCE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Existence}
Now we jump in with trying to prove existence of signed, weak energy solutions. Following closely the proof provided in \citep{vazquez2007porous}, but now lifting the assumption that $\Phi$ is strictly increasing, and rather having that it is only nondecreasing. Another useful addition to proof will be that we will also incorporate the $L^1$-contractivity result.

The proof will use an approximate problem which admits a unique solution from theorem \ref{thm:classical_existence}, and then employing the Rellich compactness argument to get a limit which hopefully will be a weak energy solution.

In the following we will assume $u_0 \in L_\Psi(\Omega)$ and $f\in L^p(\Omega)$ with $p = \frac{2d}{d+2}$ if $d \geq 3$, or for some $p>1$ if $d=1,2$\footnote{This choice of value for $p$, I can only say is a mystery for me at this moment, but it seems to be related to the Sobolev conjugate. Let's just stay curious and excited for now, shall we?}.

The main theorem reads:
\begin{theorem}
\label{thm:weak_existence}
Assume $u_0 \in L_\Psi(\Omega)$ and $f \in L^p (\Omega)$. Then the homogeneous, general porous medium equation (\ref{HDP}) has a weak solution defined in an infinite time interval, $T=\infty$. We have $u \in L^\infty(0,T:L_\Psi(\Omega))$ and $\Phi(u) \in L^2(0,T:H_0^1(\Omega))$, and the energy inequality (\ref{energy_ineq}) holds.
\end{theorem}

We will not prove this in one big chunk, but rather split it off into smaller lemmas, and as already stated, some attention will also be paid to the $L^1$-contractivity.
\subsection*{Preliminaries}
To set ourselves up in a position where we can make use of theorem \ref{thm:classical_existence}, or rather corollary \ref{cor:GPME_existence}, we need to approximate $\Phi$, $u_0$, and $f$.

We approximate $\Phi$ by the smooth approximations
\begin{align*}
\Phi_n(s) &= \left( \Phi(s) + \frac{s}{n} \right)*\eta_{\frac{1}{n}} \\
	&= \int_{\mathbb{R}}\left(\Phi(y) + \frac{y}{n}\right)\eta_{\frac{1}{n}}(s-y)dy \\
	&= \int_{\mathbb{R}}\left(\Phi(s-y) + \frac{s-y}{n}\right)\eta_{\frac{1}{n}}(y)dy
\end{align*}
where $\eta$ is the standard mollifier. We also let $u_{0,n}\in C^\infty_c(\Omega)$ be a smooth approximation of $u_0$, and converging in $L_\Psi(\Omega)$. We also ask that $|u_{0,n}| \leq $min$\{n,u_0\}$.

Apparently, also according to \citep{vazquez2007porous}, we also approximate $f$ by a sequnce of functions $f_n$ that converges to $f$ in $L^p(Q_T)$ for all\footnote{Not sure if \textit{sic}...} $p<\infty$, and is so that $|f_n| \leq f$ for all $n$.

Our first lemma then reads
\begin{lemma}
Theorem \ref{thm:weak_existence} holds with the additional assumption that $u_0$, $f$, and $\partial_t f$ are bounded, $\partial \Omega \in C^{2+\alpha}$, and $\Phi$ is locally lipschitz continuous.
\end{lemma}

\begin{proof}

\begin{description}
	\item[Setting up an approximate problem:] We fix $T<\infty$ and consider the problem
		\begin{equation}
		\label{approximate_problem}
			\begin{cases}
			\partial_t u_n &= \Delta\Phi_n(u_n) + f_n, \quad \text{ in } Q_T \\
			u_n(x,0) &= u_{0,n}(x),	\quad \text{ on } \overline{\Omega} \\
			u_n(x,t) &= 0, \quad \text{ on }  \partial \Omega \times (0,T).
			\end{cases}
		\end{equation}			
	By virtue of theorem \ref{thm:classical_existence}, this approximate problem has a unique solution $u_n \in C^{2,1}(\overline{Q}_T)$.

	\item[Control of the spatial derivative:] Now we define $w_n = \Phi_n(u_n)$, multiply the GPME by $w_n$ and integrate over $Q_T$, which results in
	\begin{equation*}
	\iint_{Q_T}w_n\partial_t u_n dxdt = \iint_{Q_T}w_n\Delta w_n dxdt + \iint_{Q_T}f_n w_n dxdt.
	\end{equation*}
	Now we do exactly as we did the section on weak energy solutions and do integration by parts on the first term on the right hand side. We also notice that $w_n \partial_t u_n = \partial_t \Psi_n(u_n)$. Thus,
	\begin{equation*}
	\int_\Omega \Psi(u_n(T))dx + \iint_{Q_T}|\nabla w_n|^2dxdt = \int_\Omega \Psi_n(u_{0,n})dx + \iint_{Q_T}f_n w_n dxdt.
	\end{equation*}
	Considering the second term on the right hand side we have
	\begin{equation*}
	\iint_{Q_T}f_n w_n dxdt \leq C\iint_{Q_T}f_n^2 dxdt + \frac{1}{2}\iint_{Q_T}|\nabla w_n|^2dxdt,
	\end{equation*}
	after exactly the same argument as we did with (\ref{using_Poincare}). After putting it back
	\begin{equation*}
	\int_\Omega \Psi_n(u_n(T))dx + \frac{1}{2}\iint_{Q_T}|\nabla w_n|^2 dxdt \leq \int_\Omega \Psi(u_{0,n})dx + C\iint_{Q_T}f_n^2 dxdt,
	\end{equation*}
	and so we're getting close to a uniform bound on $\nabla w_n$ in $L^2(Q_T)$. We notice that
	\begin{equation*}
	\iint_{Q_T}f_n^2dxdt \leq \iint_{Q_T} f^2 dxdt,
	\end{equation*}
	by our assumption on the approximation. In addition
	\begin{equation*}
	\int_\Omega \Psi_n(u_{0,n})dx = \int_\Omega \left[ \int_0^{u_{0,n}(x)} \int_\mathbb{R} \left(\Phi(y)+\frac{y}{n}\right)\eta_{\frac{1}{n}}(s-y)dy ds \right] dx,
	\end{equation*}
	and $\eta_{\frac{1}{n}}(s-\cdot)$ has support in $[s-\frac{1}{n}, s + \frac{1}{n}]$. Therefore, by the monotonicity of $\Phi$, 
	\begin{equation*}
	\int_\Omega \Psi_n(u_{0,n})dx \leq \int_\Omega \left[ \int_0^{u_{0,n}(x)}\Phi \left(s+\frac{1}{n} \right)ds + \int_0^{u_{0,n}}\frac{s+\frac{1}{n}}{n}ds \right] dx,
	\end{equation*}
	which after a variable transformation yields
	\begin{align*}
	\int_\Omega \Psi_n(u_{0,n})dx &\leq \int_\Omega \left[ \int_{\frac{1}{n} }^{u_{0,n}(x) + \frac{1}{n} } \Phi(s)ds + \frac{u_{0,n}(x)^2}{2n} + \frac{u_{0,n}(x)}{n^2} \right] dx \\
	&\leq \int_\Omega \left[ \Psi\left(u_{0,n} +\frac{1}{n}\right) - \Psi\left(\frac{1}{n}\right) + \frac{1}{2n}\left(u_{0,n} + \frac{1}{n}\right)^2 \right] dx.
	\end{align*}
	Using now that $|u_{0,n}| \leq u_0$ and that $\Omega$ is a bounded set, we get a bound on this integral independent of $n$. And so $\iint_{Q_T} |\nabla w_n|^2 dxdt$ has a bound independent of $n$.

	\item[Control of the temporal derivative:] Now we let $\zeta = \zeta(t)$ be a $C^1([0,T],[0,1])$ cutoff function so that $\zeta(0) = \zeta(T) = 1$. If we multiply (\ref{approximate_problem}) by $\zeta \partial_tw_n$ and integrate over $Q_T$ we get
	\begin{equation}
	\label{temporal_bound_start}
	\iint_{Q_T}\zeta \Phi'_n(u_n)|\partial_t u_n|^2 dxdt = \iint_{Q_T} \left\{ \zeta \partial_t w_n \Delta w_n + \zeta \partial_t w_n f_n \right\} dxdt.
	\end{equation}
	Let's start by considering the first term on the right hand side. By the product rule for the divergence
	\begin{equation*}
	\int_0^T \zeta \int_\Omega \partial_t w_n \Delta w_n dxdt = \int_0^T \zeta \int_\Omega \left\{ \nabla \cdot (\partial_t w_n \nabla w_n) - (\partial_t \nabla w_n) \cdot \nabla w_n \right\} dx dt.
	\end{equation*}
	So after using the divergence theorem together with the homogeneous boundary data, as well as noticing that $(\partial_t \nabla w_n) \cdot \nabla w_n = \frac{1}{2}\partial_t |\nabla w_n|^2$, we see that
	\begin{equation*}
	\int_0^T \zeta \int_\Omega \partial_t w_n \Delta w_n dxdt = -\iint_{Q_T}\frac{\zeta}{2}\partial_t |\nabla w_n|^2 dxdt.
	\end{equation*}
	Integrating by parts in the temporal dimension, we end up with
	\begin{equation*}
	\int_0^T \zeta \int_\Omega \partial_t w_n \Delta w_n dxdt = \iint_{Q_T} \frac{\partial_t \zeta}{2} |\nabla w_n|^2 dxdt - \frac{1}{2}\int_\Omega \left\{ |\nabla w_n(T)|^2 - |\nabla w_n(0)|^2 \right\} dx.
	\end{equation*}
	A similar inetgration by parts in the temporal dimension of the last term on the right hand side in (\ref{temporal_bound_start}) yields
	\begin{equation*}
	\iint_{Q_T} \zeta f_n \partial_t w_n dxdt = -\iint_{Q_T} \partial_t (\zeta f_n) w_n dxdt + \int_\Omega \left\{ f_n(T)w_n(T) - f_n(0)w_n(0) \right \} dx.
	\end{equation*}
	For the last term in this integral we can do the same trick as we did previously for the energy estimate. We use Cauchy-Schwarz and the inequality $\sqrt{ab} \leq \frac{c^2}{4}a + \frac{1}{c^2}b$. Then, after using a Poincaré inequality we can choose $c$ so that the integrals of $|\nabla w_n|^2$ at the temporal endpoints are cancelled. In toto we end up with
	\begin{align*}
	\iint_{Q_T} \zeta \Phi'_n(u_n) |\partial_t u_n|^2 dxdt &\leq \iint_{Q_T} \left\{ \frac{\partial_t \zeta}{2}|\nabla w_n|^2 - \partial_t(\zeta f_n) w_n \right\} dxdt \\
	&+ \frac{C}{2}\int_\Omega \left\{ f_n^2(T) - f_n^2(0) \right\} dx	\\
	&= \iint_{Q_T} \left\{ \frac{\partial_t \zeta}{2}|\nabla w_n|^2 - \partial_t(\zeta f_n) w_n \right\} dxdt \\
	&+ C \iint_{Q_T} f_n\partial_t f_n dxdt.
	\end{align*}
	Thus, by our assumption that $f$ and $\partial_t f$ are bounded, and our previous energy bound, the left hand side is bounded independent of $n$.

	In particular, we can choose $\zeta = 1$\footnote{Yikes, sorry about the kind of a fortiori argument, those are always a bit of a headache.}, and we get a bound on $\Phi'_n(u_n)|\partial_t u_n|^2$ in $L^1(Q_T)$. See now that $|\partial_t w_n	|^2 = \Phi'_n(u_n)^2|\partial_t u_n|^2$. And so, beacuse $\Phi'(u)$ is bounded for all $|u| \leq C$, we have that $\Phi'_n(u_n)$ is uniformly bounded, our latest integral bound gives us that $\partial_t w_n$ is uniformly bounded in $L^2(Q_T)$.

	\item[Using a compactness argument:] We can conclude that $\{w_n\}_{n \in \mathbb{N}}$ is a bounded sequence in $H^1(Q_T)$, so by theorem \ref{thm:Rellich} it has convergent subsequence $\{w_{n_j}\}_{j \in \mathbb{N}}$ that converges to some $w$ in $L^2(Q_T)$.

	\item[Convergence of $u_n$:] Recall that we defined $w_n = \Phi_n(u_n)$
\end{description}

\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%				APPENDICES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\begin{appendices}
%%%%%%%%%%%%%%%%%%%%%%%%%%%
%		A
%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Notation and important spaces}
Throughout this note $\Omega \subset \mathbb{R}^d$ will be a bounded domain, and $\partial \Omega$ will be its boundary. Furthermore we define $Q_T = \{ (x,t): x \in \Omega, t\in [0,T] \}$, and $\Gamma_T$ is the parabolic boundary, i.e. $\Gamma_T = \{(x,t): x \in \partial \Omega, t \in [0,T]\} \bigcup \{(x,t): x \in \Omega, t=0\}$.

\subsection{Function spaces}
Most, if not all, of the notation concerning function spaces will follow the notation in \citep{evans} (see in particular Ch. 5).

To start off $C^k(U)$ is the set of all functions that are $k$ times continuously differentiable in $U \subset \mathbb{R}^n$. A norm for continuous functions is
\begin{equation*}
	||u||_{C(\bar{U})} = \underset{x\in U}{\text{sup}}|u(x)|.
\end{equation*}
Let now $0 < \gamma \leq 1$.
\begin{mydef}[Hölder continuous]
The $\gamma^{th}$-Hölder seminorm of $u: U \to \mathbb{R}$ is
	\begin{equation*}
	|u|_{C^{0,\gamma}(\bar{U}))} = \underset{\underset{x \neq y}{x,y \in U}}{\text{sup}} \left\{\frac{|u(x)-u(y)|}{|x-y|^\gamma} \right\}
	\end{equation*}
and the $\gamma^{th}$-Hölder norm is
\begin{equation*}
||u||_{C^{0,\gamma}} = ||u||_{C(\bar{U})} + |u|_{C^{0,\gamma}}.
\end{equation*}
\end{mydef}
We continue by defining the Hölder spaces:
\begin{mydef}[Hölder space]
The Hölder space $C^{k,\gamma}(\bar{U})$ consists of all functions $u \in C^k(\bar{U})$ for which the norm
\begin{equation*}
 ||u||_{C^{k,\gamma}(\bar{U})} = \sum_{|\alpha| \leq k}||D^\alpha u||_{C(\bar{U})} + \sum_{|\alpha| = k} |D^\alpha u|_{C^{0,\gamma}(\bar{U})}
\end{equation*}
is finite.
\end{mydef}
%%%%%%%%%%%%%%%%%%
%		B
%%%%%%%%%%%%%%%%%%
\section{Passing derivative inside the integral}
\label{sec:derivative}

We are now concern with whether
\begin{equation}
\label{diff_int}
	\nabla \int_t^T \varphi(x,s)ds = \int_t^T \nabla\varphi(x,s)ds
\end{equation}
holds or not for $\varphi \in L^2(0,T:H_0^1 (\Omega))$. 
For simplicity we now only consider one spatial dimension.

To begin with, assume that $\psi, \eta \in C^{\infty}_c$, then we have that
\begin{align*}
\int_{\Omega}\psi \partial_x\left(\int_t^T \eta(x,s) ds\right)dx &= \int_{\Omega}\psi \lim_{h \to 0}\left(\int_t^T \frac{\eta(x+h,s)-\eta(x,s)}{h} ds\right)dx\\
&= \int_{\Omega}\psi \lim_{h \to 0}\left(\int_t^T \eta_h(x,s) ds\right)dx.
\end{align*}

Assuming that $\partial_x \eta(x,s)$ is bounded, we can use the mean value theorem to state that $|\eta_h| \leq  ||\partial_x \eta||_\infty$. We also have that $\eta_h(x,s) \to \partial_x \eta(x,s)$ pointwise, and so we're in a position to use the Dominated Convergence theorem to conclude that
\begin{equation}
\label{results_diff_first}
\int_{\Omega}\psi \partial_x\left(\int_t^T \eta(x,s) ds\right)dx = \int_{\Omega}\psi \left(\int_t^T \partial_x\eta(x,s) ds\right)dx.
\end{equation}

To extend the result to $L^2(0,T:H_0^1 (\Omega))$ we use that $C_c^\infty$ is dense in this space. So let $\{\psi_n\},\{\varphi_n\}$ be two sequence converging to $\psi$ and $\varphi$ in $L^2(\Omega)$ and $L^2(0,T:H_0^1(\Omega))$ respectively. Considering the left hand side of (\ref{results_diff_first}), we get
\begin{align*}
&\left|\int_{\Omega}\psi \partial_x \left( \int_t^T\varphi ds \right)dx - \int_{\Omega}\psi_n \partial_x \left( \int_t^T \varphi_n ds \right) dx \right| \\
&\leq  \left| \int_{\Omega} \psi \partial_x \left(\int_t^T \varphi - \varphi_n ds\right)dx \right| + \left| \int_{\Omega}(\psi - \psi_n)\partial_x\left(\int_t^T \varphi_n ds\right) dx \right| \\
&\leq ||\psi||_{L^2(H_0^1)}|(\varphi -\varphi_n)|_{L^2(H_0^1)} + ||\psi-\psi_n||_{L^2}| \varphi_n |_{L^2(H_0^1)} \\
&\leq ||\psi||_{L^2(H_0^1)}||(\varphi -\varphi_n)||_{L^2(H_0^1)} + ||\psi-\psi_n||_{L^2}|| \varphi_n ||_{L^2(H_0^1)} \\
&\to 0,
\end{align*}
after using the triangle- and Cauchy-Schwarz' inequality as well as the assumptions on convergence.

An exactly similar approach can be done to the right hand side, and so passing the limit in (\ref{results_diff_first}) we get
\begin{equation}
\label{result_diff_final}
\int_\Omega \psi \partial_x\left(\int_t^T \varphi(x,s)ds\right)dx = \int_\Omega \psi \left( \int_t^T \partial_x \varphi(x,s)ds \right) dx.
\end{equation}

To get the more general (\ref{diff_int}), we simply apply this result for each individual spatial dimension while keeping the others fixed.
%%%%%%%%%%%%%%%%%%%
%		C
%%%%%%%%%%%%%%%%%%%
\section{Some major theorems}
We now consider the problem of determining a function $u = u(x,t)$ in $\overline{Q}_T$ so that
\begin{equation}
\label{lady_parabolic_eq}
	u_t - \sum_{i=1}^d\frac{d}{dx_i}a_i(x,t,u,\nabla u) + a(x,t,u,\nabla u) = 0
\end{equation}
in $Q_T$ and on $\Gamma_T$
\begin{equation}
\label{lady_bc}
	u\big|_{\Gamma_T} = \psi\big|_{\Gamma_T}
\end{equation}
for some known function $\psi$.

Furthermore, we define
\begin{equation*}
A(x,t,u,\nabla u) = a(x,t,u,\nabla u) - \sum_{i=1}^d \left[ \frac{\partial a_i(x,t,u,\nabla u)}{\partial u} + \frac{\partial a_i(x,t,u,\nabla u)}{\partial x_i} \right]
\end{equation*}
Immediately below is an existence theorem presented on this type of problem taken
verbatim from \citep[ Theorem 6.1, p. 452]{ladyzhenskaya}.

\begin{theorem}[Existence of classical solutions]
\label{thm:classical_existence}
Suppose that the following conditions hold.
\begin{enumerate}[a)]
	\item For $(x,t) \in \overline{Q}_T$ and either
	\begin{equation}
		\label{exist_cond1}
		\frac{\partial a_i(x,t,u,p)}{\partial p_j}\xi_i \xi_j \Bigg|_{p=0} \geq 0, \quad A(x,t,u,0)u \geq -b_1 u^2 - b2
	\end{equation}
	or
	\begin{equation}
	\label{exist_cond2}
	A(x,t,u,0)u \geq -\Phi(|u|)|u| - b_2, \quad \int_0^\infty \frac{d\tau}{\Phi(\tau)}= \infty, \quad \Phi > 0
	\end{equation}
	for arbitrary $u$ and $\xi = (\xi_1, \xi_2, \ldots, \xi_d)$.

	\item For $(x,t) \in \overline{Q}_T$, $|u| \leq M$ and arbitrary $p$ the functions $a_i(x,t,u,p)$ and $a(x,t,u,p)$ are continuous, the $a_i(x,t,u,p)$ are differentiable with respect to $x$, $u$ and $p$ and the following inequalities are satisfied
	\begin{equation}
	\label{exist_cond3}
		\begin{cases}
			\nu\xi^2 \leq \sum_{i,j=1}^d\frac{\partial a_i(x,t,u,p)}{\partial p_j}\xi_i \xi_j \leq \mu \xi^2 \\
			\sum_{i=1}^d \left(|a_i| + |\frac{\partial a_i}{\partial u}|\right)(1+|p|) + \sum_{i,j=1}^d |\frac{\partial a_i}{\partial x_j}| + |a| \leq \mu (1 + |p|)^2. 		
		\end{cases}
	\end{equation}
	where $\nu$ and $\mu$ are positive numbers.

	\item For $(x,t) \in \overline{Q}_T$, $|u| \leq M$ and $|p| \leq M_1$, the functions $a_i$, $a$, $\frac{\partial a_i}{\partial p_j}$, $\frac{\partial a_i}{\partial u}$, and $\frac{\partial a_i}{\partial x_i}$ are continuous functions satisfying a Hölder condition in $x$, $t$, $u$ and $p$ with exponents $\beta$, $\frac{\beta}{2}$, $\beta$ and $\beta$ respectively.

	\item For $(x,t) \in \overline{Q}_T$, $|u| \leq M$ and $|p| \leq M_1$ the function $a(x,t,u,p)$ has the partial derivatives $\frac{\partial a}{\partial p_j}$ and $\frac{\partial a}{\partial u}$ and the functions $a_i$ and $a$ satisfy
	\begin{equation}
	\label{exist_cond4}
	\begin{split}
		&\Bigg| \frac{\partial a_i(x,t,u,p)}{\partial u}, \frac{a_i(x,t+h,u,p) - 		a_i(x,t,u,p)}{h}, \\
		&\frac{\partial a}{\partial p}, \frac{\partial a}{\partial u}, \frac{a(x,t+h,u,p) - a(x,t,u,p)}{h} \Bigg| \leq \varphi(x,t)
	\end{split}.
	\end{equation}

	\item $\psi(x,t) \in H^{2+\beta, 1+\beta/2}(\overline{Q}_T)$ and satisfy
	\begin{equation}
	\label{exist_cond5}
	\psi_t - \frac{d}{dx_i}a_i(x,t,\psi,\psi_x) + a(x,t,\psi,\psi_x) = 0.
	\end{equation}
	on $\partial \Omega \times \{t=0\}$.

	\item $\partial \Omega \in C^{2+\beta}$\footnote{I can't seem to find a definition of this sort of boundary in \citep{ladyzhenskaya}. However, \citep{evans} has something vaguely familiar, but here integer values of the exponent is assumed.}.
\end{enumerate}

Under these conditions there exists a unique solution of problem (\ref{lady_parabolic_eq}) with boundary condition (\ref{lady_bc}) from the class $H^{2+\beta, 1+\beta/2}(\overline{Q}_T)$. Moreover, this solution has derivatives $u_{xt}$ from $L^2(\overline{Q}_T)$.
\end{theorem}

\setcounter{rem}{0}
\begin{rem}
\label{rem:lady_notation}
By the function space $H^{l, l/2}$, Ladyzhenskaya et. al means the Banach space of functions $u$ that are continuous in $\bar{Q}_T$. So are all partial derivatives of the form $\partial_t^r D^\alpha u$, with $2r+|\alpha| <l$. The norm, they define
\begin{equation*}
	|u|_{Q_T}^{(l)} = \langle u \rangle^{(l)}_{Q_T} + \sum_{j=0}^{[l]}\langle u \rangle^{(j)}_{Q_T},
\end{equation*}
with
\begin{align*}
\langle u \rangle^{(j)}_{Q_T} &= \sum_{2r+|\alpha|=j} ||\partial_t^r D^\alpha u||_{C(\bar{Q}_T)} \\
\langle u \rangle^{(l)}_{Q_T} &= \langle u \rangle^{(l)}_{x,Q_T} + \langle u \rangle^{(l/2)}_{t,Q_T} \\
\langle u \rangle^{(l)}_{x,Q_T} &= \sum_{2r+|\alpha|=[l]} |\partial_t^r D^{\alpha}u|_{C^{0,l-[l]}(\bar{Q}_T)} \\
 \langle u \rangle^{(l/2)}_{t,Q_T} &= \sum_{0 < l - 2r -\alpha < 2} |\partial_t^r D^\alpha u|_{C^{0,(l-2r-s)/2}(\bar{Q}_T)}
\end{align*}
with the Hölder semi-norm with respect to $x$ and $t$ respectively in the last two definitions\footnote{Yikes.}.
\end{rem}

\begin{rem}
The  $\varphi$ in (\ref{exist_cond4}) is required to satisfy
\begin{equation*}
\left( \int_0^T \left(\int_\Omega |\varphi(x,t)|^q dx\right)^\frac{r}{q} dt \right)^\frac{1}{r} < \infty,
\end{equation*}
i.e. $\varphi \in L^r(0,T: L^q(\Omega))$, where $q$ and $r$ satisfy
\begin{equation*}
\frac{1}{r} + \frac{d}{2q} = 1 - x_1
\end{equation*}
and
\begin{equation*}
	\begin{cases}
		q\in \left[\frac{d}{2(1-x_1)},\infty\right], &r\in \left[\frac{1}{1-x_1},\infty \right], 0<x_1<1, \quad \text{ for } d \geq 2 \\
		q \in [1, \infty], &r \in \left[\frac{1}{1-x_1},\frac{2}{1-2x_1}\right], 0<x_1<\frac{1}{2}, \quad \text{ for } n=1
	\end{cases}
\end{equation*}
\end{rem}

\subsection{This existence result applied to the GPME}
The general porous medium equation we are concerned with can be written
\begin{equation*}
u_t - \nabla \cdot (\nabla \Phi(u) ) - f(x,t) = 0,
\end{equation*}
which is of the form (\ref{lady_parabolic_eq}) with
\begin{equation}
a_i(x,t,u,\nabla u) = \Phi'(u) u_{x_i}, \quad a(x,t,u,\nabla u) = -f(x,t).
\end{equation}

Let's go trhough each condition in theorem \ref{thm:classical_existence} to see what assumptions we must make to have a classical solution.

\begin{enumerate}[a)]
	\item We consider fullfilling (\ref{exist_cond1}). We have
	\begin{equation*}
		\frac{\partial a_i(x,t,u,p)}{\partial p_j} = \delta_{ij} \Phi'(u),
	\end{equation*}
	and so, since $\Phi'(u) \geq 0$ is assumed,
	\begin{equation*}
		\frac{\partial a_i(x,t,u,p)}{\partial p_j}\Bigg|_{p=0}\xi_i\xi_j = \Phi'(u)\xi_i^2 \geq 0.
	\end{equation*}

	In the general porous medium equation
	\begin{equation*}
		A(x,t,u,p) = -f(x,t) - \sum_{i=1}^d \Phi''(u)p_i,
	\end{equation*}
	which means $A(x,t,u,0) = -f(x,t)$ (Also see that we need $\Phi \in C^2(\mathbb{R})$). Thus, according to (\ref{exist_cond1}) we need constants $b_1$ and $b_2$ so that
	\begin{equation*}
		f(x,t)u \leq b_1 u^2 + b_2.
	\end{equation*}
	We are able to choose $b_1$ and $b_2$ if, for example, $f(x,t)$ is bounded in $\overline{Q}_T$.

	\item We need $a_i(x,t,u,p)$ and $a(x,t,u,p)$ to continuous. This translates to requiring that $\Phi'(u)$ and $f(x,t)$ are continuous. In addition $a_i(x,t,u,p)$ needs to be differentiable with respect to $x$, $t$, $u$ and $p$, which in our case means that $\Phi''(u)$ needs to be continuous.

	Considering the first inequality in (\ref{exist_cond3}) we have
	\begin{equation*}
		\sum_{i,j=1}^d \frac{\partial a_i(x,t,u,p)}{\partial p_j} \xi_i \xi_j = \sum_{i=1}^d \Phi'(u)\xi_i^2 =\Phi'(u)\xi^2.
	\end{equation*}
	Assuming that $\Phi'(u)$ is strictly positive and continuous for $|u|\leq M$ we can set
	\begin{equation}
	0 < \nu = \text{inf}_{|u| \leq M}\Phi'(u) < \text{sup}_{|u| \leq M}\Phi'(u) \leq \mu.
	\end{equation}

	After establishing that
	\begin{equation*}
		\begin{cases}
			a_i(x,t,u,p) &= \Phi'(u)p_i \\
			\frac{\partial a_i(x,t,u,p)}{\partial u} &= \Phi''(u)p_i \\
			\frac{\partial a_i(x,t,u,p)}{\partial x_j} &= 0,
		\end{cases}	
	\end{equation*}
	we put it into the left hand side of the second inequality of (\ref{exist_cond3}) as
	\begin{align*}
	&\sum_{i=1}^d\left( |a_i| + \left| \frac{\partial a_i}{\partial u}\right| \right)(1+|p|) + \sum_{i,j}^d\left| \frac{\partial a_i}{\partial x_j} \right| + |a| \\
	&= \sum_{i=1}^d\left( |\Phi'(u)| + |\Phi''(u)| \right)|p_i|(1+|p|) + |f(x,t)| \\
	&= (|\Phi'(u)| + |\Phi''(u)|)(1+|p|)\sum_{i=1}^d|p_i| + |f(x,t)|.
	\end{align*}
	By the equivalence of $l_p$-norms in finite dimensions $\sum_i |p_i| \leq C|p|$. Furthermore, we assume $\Phi''(u)$ to be continuous for $|u|\leq M$, then
	\begin{equation*}
	(|\Phi'(u)| + |\Phi''(u)|)(1+|p|)\sum_{i=1}^d|p_i| + |f(x,t)| \leq K(|p| + |p|^2) + |f(x,t)|,
	\end{equation*}
	where $K = C\cdot \text{sup}_{|u|\leq M}(|\Phi'(u)| + |\Phi''(u)|)$. Now we are almost done. All that remains is noticing that
	\begin{align*}
	K(|p| + |p|^2) + |f(x,t)| &\leq K(|p| + |p|^2) + ||f||_{\infty}(1+|p|) \\
	&\leq \text{max}\{ K, ||f||_\infty \} (1+|p|)^2.
	\end{align*}
	So we can set $\mu = \text{max}\{K, ||f||_\infty, \text{sup}_{|u|\leq M}\Phi'(u) \}$, and both inequalities in (\ref{exist_cond3}) are satisfied.

	\item To reiterate, we need the functions
	\begin{equation*}
		\begin{cases}
			a_i(x,t,u,p) &= \Phi'(u)p_i \\
			a(x,t,u,p) &= -f(x,t) \\
			\frac{\partial a_i(x,t,u,p)}{\partial p_j} &= \delta_{ij}\Phi'(u) \\
			\frac{\partial a_i(x,t,u,p)}{\partial u} &= \Phi''(u)p_i \\
			\frac{\partial a_i(x,t,u,p)}{\partial x_j} &= 0
		\end{cases}
	\end{equation*}
	to be Hölder continuous in $x$, $t$, $u$ and $p$ with exponents $\beta$, $\frac{\beta}{2}$, $\beta$ and $\beta$ respectively. If we choose $\beta = 1$ (i.e. Lipschitz continuous for all variables except $t$) the only additional assumption we need to make on $\Phi$ is that $\Phi^{(3)}(u)$ is bounded for $|u| \leq M$. The additional assumptions we need to make on $f$ is that it needs to be Lipschitz continuous with respect to $x$ and that
	\begin{equation*}
	|f(x,t) - f(x,s)| \leq C|t-s|^\frac{1}{2}
	\end{equation*}
	for some constant $C$. Note that for bounded domains, Lipschitz continuity implies Hölder continuity with lower exponent. So we could make the kind of a fortiori assumption that $f$ is Lipschitz continuous with respect to time. Or even that $f_t(x,t)$ is bounded.

	\item For this condition we see that it seems even more natural to impose that $f$ is Lipschitz continuous in $t$, and that the domain is bounded. I.e. that $T<\infty$ and $\Omega$ is bounded. Then all the previous assumptions we've made on $\Phi$ and $f$ will ensure that this condition holds.

	\item We of course have
	\begin{equation*}
		\begin{cases}
		\Psi(x,t) &= 0, \quad (x,t) \in \partial \Omega \times [0,T] \\
		\Psi(x,0) &= u_0(x), \quad x\in \Omega 
		\end{cases}
	\end{equation*}
	and so this condition isn't too hard to handle\footnote{I think...}. First off, we should require that $u_0$ vanish at the boundary, and that it is sufficiently smooth. To be more precise, $u_0$ should be in $C^2_c(\Omega)$. The condition (\ref{exist_cond5}) holds trivially, because  $\Psi\big|_{\partial \Omega \times \{t=0\}} = 0$.

	\item In this condition we can do nothing more than assume sufficient smoothness of the boundary.
\end{enumerate}

It seems only appropriate that we summarise this somehow. A nice little corollary tailored for our purposes is just what we need now.

\begin{cor}
\label{cor:GPME_existence}
Assume the following conditions are satisfied:
	\begin{enumerate}[a)]
		\item 	$\Phi: \mathbb{R} \to \mathbb{R}$ is in $C^2(\mathbb{R})$, strictly increasing, and $\Phi$ and its subsequent derivatives are all Lipschitz.

		\item $f = f(x,t)$ is $C^1(Q_T)$, bounded and Lipschitz in both its variables.

		\item $u_0$ is $C^2(\Omega)$ and $u_0$ and its derivatives vanish near the boundary $\partial \Omega$.

		\item $\partial \Omega$ is $C^{2+\beta}$-smooth.

	\end{enumerate}
Then (\ref{HDP}) has a unique solution $u\in C^{2,1}(\bar{Q}_T)$ for $T<\infty$. Moreover, this solution has derivatives $u_{xt}$ in $L^2(Q_T)$.
\end{cor}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%		COMPACTNESS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\section{Compactness argument}
Here I will simply state the Rellich-Kondrachov Compactness theorem as it is presented in \citep{evans}, mostly for my own ease of reference. For more details, please cf. Chapter 5 ibid.

\begin{mydef}
Let $X$ and $Y$ be Banach spaces with $X \subset Y$. We say that $X$ is \textbf{compactly embedded} in $Y$, written
\begin{equation*}
X \subset \subset Y,
\end{equation*}
provided 
\begin{enumerate}[i)]
	\item $||u||_Y \leq ||u||_X$ for every $u \in X$, for some constant $C$, and

	\item each bounded sequence in $X$ is precompact in $Y$.
\end{enumerate}
\end{mydef}

With this definition in mind we have the following theorem:
\begin{theorem}[Rellich-Kondrachov Compactness Theorem]
\label{thm:Rellich}
Assume $U$ is an open, bounded subset of $\mathbb{R}^n$ and $\partial U$ is $C^1$.
Suppose $1\leq p < n$. Then
\begin{equation}
\label{Rellich}
W^{1,p}(U) \subset \subset L^q(U)
\end{equation}
for each $1\leq q < p^*$, where $p^* = \frac{np}{n-p}$ is the Sobolev conjugate of $p$.
\end{theorem}
\setcounter{rem}{0}
\begin{rem}
If $p=2$, the we have
\begin{equation*}
H^1(U) \subset \subset L^q(U)
\end{equation*}
for each $1\leq q < p^*$.
\end{rem}
\begin{rem}
Since $p^* > p$ the result is valid for $q=p$. This actually holds for all $1\leq p \leq \infty$, even when $p > n$.
\end{rem}
\begin{rem}
We also have 
\begin{equation*}
W^{1,p}_0(U) \subset \subset L^p(U)
\end{equation*}
even when the assumption that $\partial U$ is $C^1$ is lifted.
\end{rem}

\end{appendices}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%				BIBLIOGRAPHY
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\bibliographystyle{plainnat}
\bibliography{project_bibliography.bib}
\end{document}